The Ontological Modal Collapse \\ as a Collapse of the Square of Opposition

### 2014-09-10 Papers/Universalis/ModalCollapse.tex
Remove apostrophe from variant MC names

A crucial step of most ontological arguments is the claim that if
God's existence is possible, then it is necessary.  This is Lemma
\textbf{L2} in G\"odel's proof.  In the modal square of opposition
(Fig. \ref{fig:square}), this is an unusual situation in which the
\textbf{I} corner must imply and entail the \textbf{A} corner, 
in the particular case when $\phi$ is $\exq x G(x)$.  
G\"odel's proof shows that his axioms are indeed 
strong enough to invert the direction of
entailment for this choice of $\phi$.  
This observation, however, immediately leads 
to the question whether the axioms are eventually 
even strong enough to enable the inverted entailment
for any arbitrary sentence $\phi$.  
That is essentially the question asked by 
Sobel \cite{Sobel1987}, and his proof of the modal collapse
(\textbf{MC}, cf. Fig.~\ref{fig:collapse}) 
provides an affirmative answer. It is possible to show
that this form of the modal collapse entails (in modal logic
\textbf{\emph{K}}) a collapse of the modal square (\textbf{MC'}),
causing the subcontraries to entail 
(and even imply) their respective
contraries. Normally, as shown in Fig. \ref{fig:square}, 
in the modal square of opposition only the 
other direction of entailment holds: the
contraries entail their subcontraries, 
assuming the \emph{modal existential import} \textbf{ExImp}
\cite{WhatToCiteHere}.
\footnote{C.: We need a reference here; I am still
  searching for good papers; this here is also interesting:
  \url{http://www.tandfonline.com/doi/pdf/10.1080/01445340.2013.764962}}

Moreover, in any modal logic where the axiom \textbf{T} holds
(i.e. where the accessibility relation is reflexive), 
even a total collapse of the modalities (\textbf{MC''}) 
is entailed by \textbf{MC}.
Interestingly, under this stronger form of modal
collapse, the contraries entail their subcontraries even 
without the existential import.

@@

\begin{figure}[t]
\noindent \framebox[\columnwidth][r]{
\begin{minipage}{.94\columnwidth}\small
\begin{itemize}
\item[\textbf{MC}] Everything that is the case is so necessarily:
  $\hol{\allq \phi [\phi \imp \nec \phi]}$ \\
\item[\textbf{MC'}] Everything that is possible is necessary:
  $\hol{\allq \phi [\pos \phi \imp \nec \phi]}$  \\
\item[\textbf{T}] Everything that is necessary is the case:
  $\hol{\allq \phi [\nec \phi \imp \phi]}$ \\
\item[\textbf{ExImp}] (Modal Existential Import):
  $\hol{\pos \top}$ \\
\item[\textbf{AI}] Everything that is necessary is possible:
  $\hol{\allq \phi [\nec \phi \imp \pos \phi]}$ \\
\item[\textbf{MC''}] Modalities collapse completely:
  $\hol{\allq \phi [(\phi \biimp \nec \phi) \wedge (\pos \phi \biimp \nec \phi)]}$
\end{itemize}
\end{minipage}
} \vskip-.5em
\caption{Modal Collapse}
\label{fig:collapse} 
\end{figure}

>>>>

A crucial step of most ontological arguments is the claim that if
God's existence is possible, then it is necessary.  This is Lemma
\textbf{L2} in G\"odel's proof.  In the modal square of opposition
(Fig. \ref{fig:square}), this is an unusual situation in which the
\textbf{I} corner must imply and entail the \textbf{A} corner, 
in the particular case when $\phi$ is $\exq x G(x)$.  
G\"odel's proof shows that his axioms are indeed 
strong enough to invert the direction of
entailment for this choice of $\phi$.  
This observation, however, immediately leads 
to the question whether the axioms are eventually 
even strong enough to enable the inverted entailment
for any arbitrary sentence $\phi$.  
That is essentially the question asked by 
Sobel \cite{Sobel1987}, and his proof of the modal collapse
(\textbf{MC}, cf. Fig.~\ref{fig:collapse}) 
provides an affirmative answer. It is possible to show
that this form of the modal collapse entails (in modal logic
\textbf{\emph{K}}) a collapse of the modal square (\textbf{MCs}),
causing the subcontraries to entail 
(and even imply) their respective
contraries. Normally, as shown in Fig. \ref{fig:square}, 
in the modal square of opposition only the 
other direction of entailment holds: the
contraries entail their subcontraries, 
assuming the \emph{modal existential import} \textbf{ExImp}
\cite{WhatToCiteHere}.
\footnote{C.: We need a reference here; I am still
  searching for good papers; this here is also interesting:
  \url{http://www.tandfonline.com/doi/pdf/10.1080/01445340.2013.764962}}

Moreover, in any modal logic where the axiom \textbf{T} holds
(i.e. where the accessibility relation is reflexive), 
even a total collapse of the modalities (\textbf{MCt}) 
is entailed by \textbf{MC}.
Interestingly, under this stronger form of modal
collapse, the contraries entail their subcontraries even 
without the existential import.

@@

\begin{figure}[t]
\noindent \framebox[\columnwidth][r]{
\begin{minipage}{.94\columnwidth}\small
\begin{itemize}
\item[\textbf{MC}] Everything that is the case is so necessarily:
  $\hol{\allq \phi [\phi \imp \nec \phi]}$ \\
\item[\textbf{MCs}] Everything that is possible is necessary:
  $\hol{\allq \phi [\pos \phi \imp \nec \phi]}$  \\
\item[\textbf{T}] Everything that is necessary is the case:
  $\hol{\allq \phi [\nec \phi \imp \phi]}$ \\
\item[\textbf{ExImp}] (Modal Existential Import):
  $\hol{\pos \top}$ \\
\item[\textbf{AI}] Everything that is necessary is possible:
  $\hol{\allq \phi [\nec \phi \imp \pos \phi]}$ \\
\item[\textbf{MCt}] Modalities collapse completely:
  $\hol{\allq \phi [(\phi \biimp \nec \phi) \wedge (\pos \phi \biimp \nec \phi)]}$
\end{itemize}
\end{minipage}
} \vskip-.5em
\caption{Modal Collapse}
\label{fig:collapse} 
\end{figure}
